Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections

### 2022-05-19 manuscript.tex
minor edits [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/5f9575bbee2059d4e92d5c821f29682c4a661e49.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/5f9575bbee2059d4e92d5c821f29682c4a661e49/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2354476139

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Visual proteomics using whole-lamella 2D template matching},
  pdfauthor={Johannes ElferichNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}

@@

\title{Visual proteomics using whole-lamella 2D template matching}
\author{}
\date{}

@@

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/2d0a9a5debbbf44b5f9d5797d38db68589012013/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/2d0a9a5debbbf44b5f9d5797d38db68589012013}{jojoelfe/deco\_lace\_template\_matching\_manuscript@2d0a9a5}
on March 30, 2022.

@@

Localization of biomolecules inside a cell is an important goal of biological
imaging. Fluorescence microscopy can localize biomolecules inside whole cells
and tissues, but its ability to count biomolecules and accuracy of the spatial
coordinates is limited by the wavelength of visible light. Cryo-electron
microscopy (cryo-EM) provides highly accurate position and orientation
information of biomolecules but is often confined to small fields of view inside
a cell, limiting biological context. In this study we use a new data-acquisition
scheme called ``Defocus-Corrected Large-Area cryo-EM'' (DeCo-LACE) to collect
high-resolution cryo-EM data over entire sections (100 -- 200 nm thick lamellae)
of neutrophil-like mouse cells, representing roughly 1\% of the total cellular
volume. We use 2D template matching (2DTM) to determine localization and orientation
of the large ribosomal subunit in these sections, detect bound small
ribosomal subunits and assign ribosomes to polysomes based on their relative orientations to each other. These data provide ``maps'' of translational activity across
sections of mammalian cells. This new high-throughput cryo-EM
data collection approach together with 2DTM will advance visual
proteomics and complement other single-cell ``omics'' techniques, such as
flow-cytometry and single-cell sequencing.

@@

A major goal in understanding cellular processes is the knowledge of the
amounts, location, interactions, and conformations of biomolecules
inside the cell. This knowledge can be obtained by approaches broadly
divided into label- and label-free techniques. In label-dependent
techniques a probe is physically attached to a molecule of interest that
is able to detected with a high signal-to-noise signal, such as a
fluorescent molecule. In label-free techniques the physical properties
of molecules themselves are used for detection. An example for this is
proteomics using mass-spectrometry {[}\protect\hyperlink{ref-tSXIKPl7}{1}{]}. The advantage
of label-free techniques is that they can provide information over
thousands of molecules, while label-techniques offer highly specific
information for a few molecules. Especially spatial information can
often only be achieved using label-dependent techniques, such as
fluorescence microscopy {[}\protect\hyperlink{ref-VBmW7Aot}{2}{]}.

Cryo-electron microscopy has the potential to directly visualize the
arrangement of atoms that compose biomolecules inside of cells, thereby
allowing label-free detection with high spatial accuracy. This has been
called ``visual proteomics'' {[}\protect\hyperlink{ref-tGQ6TSUo}{3}{]}. Since cryo-EM
requires thin samples (\textless500nm), imaging of biomolecules inside cells is
restricted to small organisms, thin regions of cells, or samples that
have been suitably thinned. Thinning can be achieved either by
mechanical sectioning {[}\protect\hyperlink{ref-g8QavfwP}{4}{]} or by
milling using a focused ion beam (FIB) {[}\protect\hyperlink{ref-16IhS1Nc4}{5}{]}. his complex workflow
leads to a low throughput of cryo-EM imaging of cells and is further
limited by the fact that at the required magnifications, typical fields
of view (FOV) are very small compared to mammalian cells, and the FOV
achieved by label-techniques such as fluorescence light microscopy. The
predominant cryo-EM technique for the localization of biomolecules of
defined size and shape inside cells is cryo-electron tomography {[}\protect\hyperlink{ref-Rksh2dxu}{6}{]}. However, the requirement of a tilt series at
every imaged location and subsequent image alignment, severely limits
the throughput for molecular localization.

An alternative approach is to identify molecules by their structural
``fingerprint'' in single projection using ``2D template-matching'' (2DTM)
{[}\protect\hyperlink{ref-Ynb3IP6I}{7},\protect\hyperlink{ref-18KGpXYPE}{8},\protect\hyperlink{ref-10bXZuF3G}{9}{]}. In this
method a 3D model of a biomolecule is used as a template to find 2D
projections that match the molecules visible in the electron
micrographs. This method requires a projection search on a fine angular
grid, and the projections are used to find local cross-correlation peaks
with the micrograph. Since the location of a biomolecule in the
z-direction causes predictable aberrations to the projection image, this
method can be used to calculate complete 3D coordinates and orientations
of a biomolecule in a cellular sample
{[}\protect\hyperlink{ref-18KGpXYPE}{8}{]}

Hematopoiesis is the process of generating the various cell types of the blood in
the bone marrow. Disregulation of the process results in diseases like leukemia.
Understanding how hematopoietic stem and progenitor cells are programmed to
diffferentiate to the appropriate cell type would be provide new insight how
hematopoiesis can be misregulated. Of special interest is the regulation of
translation during hematopoiesis. This is exemplified by the observation that
genetic defects in the ribosome machinery often leads to hematopoietic
disease{[}\protect\hyperlink{ref-gRoY21jY}{10}{]}. As such direct quantification of ribosome
location, number and conformational states could lead to new insight into
hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{11}{]}.

Here we apply 2D-template matching of ribosomes to cryo-FIB milled
neutrophil-like murine cells {[}\protect\hyperlink{ref-1B9Vt9eYu}{12}{]}. To increase the
amount of collected data and to provide unbiased sampling of the whole
lamella, we devised a new data-acquisition scheme, Defocus-corrected
large area cryo-electron microscopy (DeCo-LACE). We characterize
aberration cause by the used large beam-image shifts and highly focused
beams and find that they can be adequately corrected to enable ribosome
detection by 2DTM. The resulting data provide a description of ribosome
distribution in the whole lamellae, which represent roughly 2\% of the
cellular volume. We find highly heterogeneous density of ribosome within
the cell and can identify discrete clusters of presumably
translationally active ribosomes, by testing for the presence of the
small ribosomal subunit. The high accuracy of location and orientation
of each detected ribosome also allows us to cluster ribosome molecules
into potential polysomes. Analysis of the throughput in this method
suggests that for the foreseeable future computation will be the
bottleneck for visual proteomics.

@@

ER-HoxA9 cells were maintained in RPMI supplemented with 10\% FBS,
penicillin/streptomycin, SCF, and estrogen {[}\protect\hyperlink{ref-1B9Vt9eYu}{12}{]} at
37C and 5\% CO2. 120h prior to grid freezing cells were washed twice in PBS and
cultured in the same medium, except without estrogen. Differentiation was
verified by staining with Hoechst-dye and insepction of nuclear morphology.
Cells were then counted and diluted to 1\^{}106 cells/ml. Grids ( either 200 mesh
copper grids, with a sillicone-oxide and 2um holes with a 2um spacing or 200
mesh gold grids with a thin gold film and 2 um holes in 2um sapcing) were
glow-discharged from both sides using a \ldots{} for \ldots{} . 3.5 ul of cells suspension
was added to grids on the thin-film side and grids were automatically blotted
from the back-side using a GP2 cryoplunger (Leica) for \ldots{} s and rapidly plunged
into liquid ethane at -185C.

@@

Grids were loaded into a Acquilos 2 FIB/SEM microscope with a stage cooled to
-190C. Grids were sputter-coated with platinum for 15s at 45 mA and then coated
with a layer of platinum-precursor by openin the GIS-valve for 45s. An overview
of the grid was created by montaging SEM images and isolated cells at the center
of gridsquares were selected for FB-milling. Lamella were generated
automatically using the AutoTEM software, resulting in 6-8 um wide lamella with
150-200 um thickness as determined by FIB-imaging of the lamella edge.

@@

Grids were loaded into a Krios Titam TEM operated at 300 keV. The microscope was
setup with a cross-grating grid on the stageby setting the beam-diameter to 900
nm, resulting in the beam being completely visible in the camera. To establish
fringe-free conditions, the ``Fine eucentric'' procedure of serialEM was used to
move a square of the cross-grating grid to the eucentric position of the
microscope. The effective defocus was then set to 2 um, using the ``autofocus''
routine of serialEM. The objective focus of the microscope was changed until no
fringes were visible. The stage was then moved in Z until images had a apparent
defocus of 2 um. The differnce in stage Z-position between the eucentric and
fringe-free conditions was calculate d and noted to move other areas into
fringe-free condition.

Low magnification montages were used to find lamella and lamella that were
sufficiently thin and free of contamination were selected for automated data
collection. The corners of the lamella were manually annotated in SerialEM and
translated into Beam-Imageshift values using SerialEm calibration. A hexagonal
patter of beam-imageshift positions was calculated that covered the area between
he four corners in a serpentine way, with a sqrt(3) * 400 nm horizontal spacing
and 800 nm vertical spacing. Exposures were then taking at each position with a
30 e/A total dose. After each exposure that defocus was estimated using the
ctffind function of SerialEM and the focus for th next exposure was corrected by
the difference between the estimated focus and the desired defocus of 800 um.
Also after each exposure the deviation of the beam from the center of the camera
was measured and corrected using the ``CenterBeamFrom IMage'' command of SerialEM.

After datacollection a 20s exposure at 2250x magnification of the lamella at
200um defocus was taken for visualization purposes.

\hypertarget{data-pre-processing}{%
\subsubsection{Data pre-processing}\label{data-pre-processing}}

Movies were gain-corrected and motion-corrected using a custom version of
unblur. To avoid influence of the beam-edge on motion-correction only a quarter
of the movie in the center of the camera was considered for calculation of the
estimated motion. After movie frames were aligned and summed a mask for the
illuminated area was calculated by lowpass filtering the image at \ldots{} A,
thresholding the image at 10\% of the maximal value and then lowpass filtering
the mask at \ldots{} A. This mask was then used to replace un-illuminated area with
gaussian noise, with the same mean and standard deviation as the illuminated
area. The contrast-transfer function (CTF) was estimated using ctffind, searching
between 0.02 and 2 um defocus.

\hypertarget{template-matching}{%
\subsubsection{Template matching}\label{template-matching}}

The search template was generated from the cryo-EM structure of the mouse large
ribosomal subunit (PDB 6SWA). The \ldots{} subunit was deleted from the model and the
simulate program of the cisTEM suite was used to calculate an density map from
the atomic coordinates. The match\_template program was used to search for this
template in the preprocessed images, using 1.5 deg angular step in out-of-plane
angles and 1.0 deg in-plane. 21 defocus planes in 200nm steps centered around
the defocus estimates by ctffind were searches. Matches were defined as peaks
above a threshold calulated according to .. .(7.75 for most images).

\hypertarget{data-analysis}{%
\subsubsection{Data analysis}\label{data-analysis}}

@@

\hypertarget{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamella-of-mammalian-cells}{%
\subsubsection{2DTM detects large ribosomal subunits in cryo-FIB lamella of mammalian cells}\label{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamella-of-mammalian-cells}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}

@@

\title{Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections}
\author{}
\date{}

@@

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/5f9575bbee2059d4e92d5c821f29682c4a661e49/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/5f9575bbee2059d4e92d5c821f29682c4a661e49}{jojoelfe/deco\_lace\_template\_matching\_manuscript@5f9575b}
on May 19, 2022.

@@

Localization of biomolecules inside a cell is an important goal of biological
imaging. Fluorescence microscopy can localize biomolecules inside whole cells
and tissues, but its ability to count biomolecules and accuracy of the spatial
coordinates is limited by the wavelength of visible light. Cryo-electron
microscopy (cryo-EM) provides highly accurate position and orientation
information of biomolecules but is often confined to small fields of view inside
a cell, limiting biological context. In this study we use a new data-acquisition
scheme called ``Defocus-Corrected Large-Area cryo-EM'' (DeCo-LACE) to collect
high-resolution cryo-EM data over entire sections (100 -- 200 nm thick lamellae)
of neutrophil-like mouse cells, representing 1-2\% of the total cellular
volume. We use 2D template matching (2DTM) to determine localization and orientation
of the large ribosomal subunit in these sections. These data provide ``maps'' of translational activity across
sections of mammalian cells. This new high-throughput cryo-EM
data collection approach together with 2DTM will advance visual
proteomics and complement other single-cell ``omics'' techniques, such as
flow-cytometry and single-cell sequencing.

@@

A major goal in understanding cellular processes is the knowledge of the
amounts, location, interactions, and conformations of biomolecules
inside the cell. This knowledge can be obtained by approaches broadly
divided into label- and label-free techniques. In label-dependent
techniques a probe is physically attached to a molecule of interest that
is able to be detected by its strong signal, such as a
fluorescent molecule. In label-free techniques, the physical properties
of molecules themselves are used for detection. An example for this is
proteomics using mass-spectrometry {[}\protect\hyperlink{ref-tSXIKPl7}{1}{]}. The advantage
of label-free techniques is that they can provide information over
thousands of molecules, while label-dependent techniques offer highly specific
information for a few molecules. Especially spatial information can
often only be achieved using label-dependent techniques, such as
fluorescence microscopy {[}\protect\hyperlink{ref-VBmW7Aot}{2}{]}.

Cryo-electron microscopy (cryo-EM) has the potential to directly visualize the
arrangement of atoms that compose biomolecules inside of cells, thereby
allowing label-free detection with high spatial accuracy. This has been
called ``visual proteomics'' {[}\protect\hyperlink{ref-tGQ6TSUo}{3}{]}. Since cryo-EM
requires thin samples (\textless500nm), imaging of biomolecules inside cells is
restricted to small organisms, thin regions of cells, or samples that
have been suitably thinned. Thinning can be achieved either by
mechanical sectioning {[}\protect\hyperlink{ref-g8QavfwP}{4}{]} or by
milling using a focused ion beam (FIB) {[}\protect\hyperlink{ref-16IhS1Nc4}{5}{]}. This complex workflow
leads to a low throughput of cryo-EM imaging of cells and is further
limited by the fact that at the required magnifications, typical fields
of view (FOV) are very small compared to mammalian cells, and the FOV
achieved by label-dependent techniques such as fluorescence light microscopy. The
predominant cryo-EM technique for the localization of biomolecules of
defined size and shape inside cells is cryo-electron tomography {[}\protect\hyperlink{ref-Rksh2dxu}{6}{]}. However, the requirement of a tilt series at
every imaged location and subsequent image alignment, severely limits
the throughput for molecular localization.

An alternative approach is to identify molecules by their structural
``fingerprint'' in single projection using ``2D template-matching'' (2DTM)
{[}\protect\hyperlink{ref-Ynb3IP6I}{7},\protect\hyperlink{ref-18KGpXYPE}{8},\protect\hyperlink{ref-10bXZuF3G}{9}{]}. In this
method, a 3D model of a biomolecule is used as a template to find 2D
projections that match the molecules visible in the electron
micrographs. This method requires a projection search on a fine angular
grid, and the projections are used to find local cross-correlation peaks
with the micrograph. Since the location of a biomolecule in the
z-direction causes predictable aberrations to the projection image, this
method can be used to calculate complete 3D coordinates and orientations
of a biomolecule in a cellular sample
{[}\protect\hyperlink{ref-18KGpXYPE}{8}{]}.

Here we apply 2DTM of the ribosome large subunit (LSU) to cryo-FIB milled
neutrophil-like murine cells {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]}. We chose these
cells because genetic defects in the ribosome machinery often leads to
hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and direct quantification of
ribosome location, number and conformational states in hematopoietic cells could
lead to new insight into hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}.
To increase the amount of collected data and to provide unbiased sampling of the
whole lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test whether
aberrations caused by large beam-image shifts and highly condensed beams
deteriorate the high-resolution signal. We find that these aberrations do not
impede LSU detection by 2DTM. The resulting data provide a description of
ribosome distribution in an entire lamella, which represent 1-2\% of the cellular
volume. We find a highly heterogeneous density of ribosomes within the cell.
Analysis of the throughput in this method suggests that for the foreseeable
future computation will be the bottleneck for visual proteomics.

@@

ER-HoxA9 cells were maintained in RPMI medium supplemented with 10\% FBS,
penicillin/streptomycin, SCF, and estrogen {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} at
37 °C and 5\% CO2. 120 h prior to grid preparation, cells were washed twice in PBS
and cultured in the same medium except without estrogen. Differentiation was
verified by staining with Hoechst-dye and inspection of nuclear morphology.
Cells were then counted and diluted to \(1\cdot10^6\) cells/ml. Grids (either 200
mesh copper grids, with a sillicone-oxide and 2 µm holes with a 2 µm spacing or
200 mesh gold grids with a thin gold film and 2 µm holes in 2 µm spacing) were
glow-discharged from both sides using a 15 mA for 45 s. 3.5 µl of cell suspension
was added to grids on the thin-film side and grids were blotted from the back
side using a GP2 cryoplunger (Leica) for 8 s and rapidly plunged into liquid
ethane at -185 °C.

@@

Grids were loaded into an Aquilos 2 FIB/SEM (Thermo Fisher) instrument with a
stage cooled to -190 °C. Grids were sputter-coated with platinum for 15 s at 45 mA
and then coated with a layer of platinum-precursor by opening the GIS-valve for
45 s. An overview of the grid was created by montaging SEM images and isolated
cells at the center of gridsquares were selected for FIB-milling. Lamellae were
generated automatically using the AutoTEM software (Thermo Fisher), with the
following parameters:

\begin{itemize}
\tightlist
\item
  Milling angle: 20°
\item
  Rough milling: 3.2 µm thickness, 0.5 nA current
\item
  Medium milling: 1.8 µm thickness, 0.3 nA current, 1.0° overtilt
\item
  Fine milling: 1.0 µm tchickness, 0.1 nA current, 0.5° overtilt
\item
  Finer milling: 700 nm thickness, 0.1 nA curent, 0.2° overtilt
\item
  Polish 1: 450 nm thickness, 50 pA current
\item
  Polish 2: 200 nm thickness, 30 pA current
\end{itemize}

This resulted in 6-10 µm wide lamella with 150-250 nm thickness as determined by
FIB-imaging of the lamella edges.

@@

Grids were loaded into a Titan Krios TEM (Thermo Fisher) operated at 300 keV and
equipped with a BioQuantum energy filter (Gatan) and K3 camera (Gatan). The
microscope was aligned using a cross-grating grid on the stage. Prior to each
session, we carefully performed the ``Image/Beam'' calibration in nanoprobe. We
set the magnification to a pixel size of 1.76 Å and condensed the beam to \textasciitilde{} 900 nm
diameter, resulting in the beam being completely visible on the camera. To
establish fringe-free conditions, the ``Fine eucentric'' procedure of SerialEM {[}\protect\hyperlink{ref-19ZFerhph}{13}{]} was
used to move a square of the cross-grating grid to the eucentric position of the
microscope. The effective defocus was then set to 2 µm, using the ``autofocus''
routine of SerialEM. The objective focus of the microscope was changed until no
fringes were visible. The stage was then moved in Z until images had an apparent
defocus of 2 µm. The difference in stage Z-position between the eucentric and
fringe-free conditions was used to move other areas into fringe-free condition.

Low magnification montages were used to find lamellae and lamellae that were
sufficiently thin and free of contamination were selected for automated data
collection. Overview images of each lamella were taken at 2250x magnification
(38 Å pixel size). The corners of the lamella in the overview image were manually
annotated in SerialEM and translated into beam image-shift values using SerialEM's
calibration. A hexagonal pattern of beam image-shift positions was calculated
that covered the area between the four corners in a serpentine way, with a
\(\sqrt{3}\cdot425\) nm horizontal spacing and \(3/4\cdot 850\) nm vertical spacing. Exposures were
taken at each position with a 30 e\textsuperscript{-}/Å\(^2\) total dose. After each exposure, the
defocus was estimated using the ctffind function of SerialEM and the focus for
the next exposure was corrected by the difference between the estimated focus
and the desired defocus of 800 nm. Furthermore, after each exposure the
deviation of the beam from the center of the camera was measured and corrected
using the ``CenterBeamFromImage'' command of SerialEM.

After data collection, a 20 s exposure at 2250x magnification of the lamella at
200 μm defocus was taken for visualization purposes. A Python script implementing
this procedure is available at {[}Link to repo{]}.

\hypertarget{deco-lace-data-processing}{%
\subsubsection{DeCo-LACE data processing}\label{deco-lace-data-processing}}

An overview of the data analysis pipeline is shown in Fig.
\ref{fig:deco_lace_workflow}.

\hypertarget{pre-processing}{%
\paragraph{Pre-processing}\label{pre-processing}}

Motion-correction, dose weighting and other preprocessing as detailed below was performed using \emph{cis}TEM {[}\protect\hyperlink{ref-MUY42yxp}{14}{]}. To avoid influence of the beam-edge on motion-correction, only a quarter of the
movie in the center of the camera was considered for calculation of the
estimated motion. After movie frames were aligned and summed, a mask for the
illuminated area was calculated by lowpass filtering the image with a 100 Å
resolution cutoff, thresholding the image at 10\% of the maximal value and then lowpass
filtering the mask again with a 100 Å resolution cutoff to smooth the mask edges. This mask was
then used to select dark areas in the image and fill the pixels with
Gaussian noise, with the same mean and standard deviation as the illuminated
area. A custom version of the unblur program {[}\protect\hyperlink{ref-1G4Y94qCy}{15}{]}
implementing this procedure is available at {[}link to decolace branch{]}. During motion correction images were resampled to a pixel size of 1.5 Å. The
contrast-transfer function (CTF) was estimated using ctffind
{[}\protect\hyperlink{ref-n43f3Sqs}{16}{]}, searching between 0.2 and 2 μm defocus.

\hypertarget{dtm}{%
\paragraph{2DTM}\label{dtm}}

The search template was generated from the atomic model of the mouse LSU
(PDB 6SWA, exluding the Epb1 subunit) using the cryo-EM simulator implemented in \emph{cis}TEM
{[}\protect\hyperlink{ref-q2PSamkS}{17}{]}. The
match\_template program {[}\protect\hyperlink{ref-10bXZuF3G}{9}{]} was used to search for this
template in the movie-aligned, exposure-filtered and masked images, using a 1.5°
angular step in out-of-plane angles and a 1.0° angular step in-plane. 11 defocus
planes in 20 nm steps centered around the ctffind-determined defocus were searched. Targets
were defined as detected when their matches with the template produced peaks
with a singal-to-noise ratio (SNR) above a threshold of 7.75, which was chosen
based on the one-false-positive-per-tile criterion {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}.

\hypertarget{montage-assembly}{%
\paragraph{Montage assembly}\label{montage-assembly}}

The coordinates of each tile \(i\),
\(\mathbf{c}_{i}\) {[}2D Vector in pixels{]} were initialized using beam image-shift of the tile, \(\mathbf{b}_i\) {[}2D Vector in μm{]},
and the ISToCamera matrix \(\mathbf{IC}\), as calibrated by SerialEM:

\[\mathbf{c}_{i} = \mathbf{IC} \cdot \mathbf{b}_i\]

A list of tile pairs \(i,j\) that overlap were assembled by selecting images where
\(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to calculate the precise offset between tiles \(i\) and \(j\), \(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{18}{]} was used to calculate refined offsets . The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square minimization against \(\mathbf{r}_{i,j}\):

\[ \displaystyle{\min_{\mathbf{c}} \sum_{pairs}{(\mathbf{r}_{i,j} - (\mathbf{c}_i-\mathbf{c}_j))^2}}\]

using the scipy package {[}\protect\hyperlink{ref-8Miti2Gz}{19}{]}. The masked cross-correlation and the least-square minimization was repeated once more to arrive at the final tile alignment.

The x,y coordinates of target \(n\) detected by 2DTM in the tile \(i\),
\(\textbf{m}^\textrm{T}_{n,i}\), was transformed into the montage frame by adding
the coordinate of the tile.

\[ \textbf{m}^\textrm{M}_n = \textbf{m}^\textrm{T}_{n,i} + \textbf{c}_i\]

The z coordinate of each target was calculated as the sum of the defocus offset
for the target, the estimated defocus of the tile, and the nominal defocus of
the microscope when the tile was acquired.

Images were rendered using UCSF ChimeraX {[}\protect\hyperlink{ref-cYCuuZxG}{20}{]}. The Python scripts used for data processing are available
under {[}repolink{]}.

@@

\hypertarget{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamellae-of-mammalian-cells}{%
\subsubsection{2DTM detects large ribosomal subunits in cryo-FIB lamellae of mammalian cells}\label{dtm-detects-large-ribosomal-subunits-in-cryo-fib-lamellae-of-mammalian-cells}}

To test whether we could detect individual ribosomes in mammalian cells we
prepared cryo-lamellae of mouse neutrophil-like cells. An overview image
of a representative lamella clearly shows cellular features consistent with a
neutrophil-like phenotype, mainly a segmented nucleus and a plethora of
membrane-organelles, corresponding to the granules and secretory vesicles of
neutrophils (Fig. {[}\ref{fig:initmatching}{]}A). We then proceeded to acquire
micrographs on this lamella with a defocus of 0.5-1.0 μm, 30 e\textsuperscript{-}/Å\(^2\)/s
exposure and 1.76 Å pixel size. We manually selected multiple locations in the
lamella and acquired micrographs using standard low-dose techniques where
focusing is performed on a sacrificial area. The resulting micrographs showed
smooth bilayered membranes and no signs of crystalline ice (Fig.
{[}\ref{fig:initmatching}{]}C,D), indicating successful vitrification throughout the
lamella. Successful vitrification is facilitated by the small size (\textasciitilde8 μm diameter) of these cells.
